Assessing spatial correlations in pulsar timing arrays with the noise-marginalized optimal statistic

### 2018-04-26 manuscript/manuscript.tex
major edits, mostly to Sec. II

However, the optimal statistic gives biased results 
for PTAs with significant red noise in the individual pulsars due to the 
strong covariance between the individual red noise parameters 
and the GWB amplitude \sv{has this ever been discussed in a paper?}. 
%In this paper, we present a technique for improving the accuracy of the optimal statistic 
%by marginalizing the optimal statistic over the individual red noise parameters 
%using the posterior distributions from a full Bayesian analysis. 
%This hybrid approach produces a more accurate 
%estimate of \Agw\ and its uncertainty. 
Here we present a technique for improving the accuracy of the optimal statistic by including an additional step: 
marginalizing over the individual pulsar red noise parameters 
using the posterior distributions from a full Bayesian analysis of all the pulsars. 
This hybrid approach produces a more precise estimate of the stochastic GWB amplitude \Agw\ 
and its uncertainty, while requiring only a few minutes over the more traditional method of computation. 
%This method is more computationally intensive than previous methods 
%for computing the optimal statistic, but still requires only minutes. 
%Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
%can be used to compute the optimal statistic with any choice of the ORF, 
%allowing additional spatial correlations to be tested very quickly.
Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
can be used to compute the optimal statistic with any choice of overlap-reduction function (ORF), 
allowing for quick testing of various spatial correlations. 
This technique was used to perform the frequentist searches 
for common red signals with monopole, dipole, and Hellings-Downs spatial correlations 
in the NANOGrav 11-year dataset \citep{abb+17b}.

@@

When constructing the residuals, 
we typically fix the red noise parameters to the values that 
maximize the likelihood. 
However, this leads to a bias in the optimal statistic and SNR because 
the individual red noise parameters and the common red noise parameters 
are highly covariant. 
In this section, we compare three different techniques for computing the 
optimal statistic for PTAs with significant red noise. 
First, we fix the individual pulsars' red noise parameters to the 
maximum-likelihood values from individual pulsar analyses. 
Second, we fix the individual pulsars' red noise parameters to the 
values that jointly maximize the likelihood for a common analysis 
of all of the pulsars in our PTA 
that searches over the pulsars' red noise parameters and 
a common red process. 
For the third method, we draw values of the individual pulsars' 
red noise parameters from the posteriors generated by a Bayesian analysis 
that includes both individual red noise and a common red noise process, 
allowing us to marginalize over the individual red noise.

We used these three techniques to compute the optimal statistic 
for simulated ``NANOGrav-like'' data sets. 
These consisted of 18 MSPs with observation times up to 11 years 
and sky positions matching the 18 longest-observed pulsars in the 
NANOGrav 11-year data set \citep{abb+17}.
with the observation times increased so that the longest-observed pulsars 
had TOAs spanning over 11 years. 
We include white noise for all pulsars, plus 
red noise parametrized as a power law, $P(f) = A_\mathrm{red} f^{-\gamma}$, 
for those pulsars that show evidence of red noise 
(see Table~\ref{tab:sim} for more details). 
To these data sets we injected an isotropic stochastic GWB 
\sv{insert strength of GWB... are we going to have one value or more than one?}.
\begin{table}[htb]
	\setlength{\tabcolsep}{5pt}
	\caption{Pulsar parameters used in simulated PTAs. \sv{finish filling in this table}}
	\begin{center}
	\begin{tabular}{ccccc}
		\hline\hline
    		Pulsar	& $T_\mathrm{obs}$ (yrs) & $\sigma_w$ ($\mu\mathrm{s}$) & $A_\mathrm{red}$ & $\gamma_\mathrm{red}$ \\
		\hline
		J0030+0451 & 11.0 & & -13.93 & 3.56 \\
		J0613$-$0200 & 11.0 &  & -13.14 & 1.22 \\
		J1012+5307 & 11.0 &  & -12.79 & 1.51 \\
		J1024$-$0719 & 6.0 & & $-$ & $-$ \\
		J1455$-$3330 & 11.0 & & $-$ & $-$ \\
		J1600$-$3053 & 8.0 & & $-$ & $-$ \\
		J1614$-$2230 & 7.0 & & $-$ & $-$ \\
		J1640+2224 & 11.0 & & $-$ & $-$ \\
		J1713+0747 & 11.0 & & -14.14 & 1.58 \\
		J1741+1351 & 6.0 & & $-$ & $-$ \\
		J1744$-$1134 & 11.0 & & $-$ & $-$ \\
		B1855+09 & 11.0 & & -13.75 & 3.54 \\
		J1853+1303 & 7.0 & & $-$ & $-$ \\
		J1909$-$3744 & 11.0 &  & -13.84 & 1.74  \\
		J1918$-$0642 & 11.0 & & $-$ & $-$ \\
		J2010$-$1323 & 7.0 & & $-$ & $-$ \\
		J2145$-$0750 & 11.0 &  & -12.69 & 1.30 \\
		J2317+1439 & 11.0 & & $-$ & $-$ \\
    		\hline\hline
	\end{tabular}
	\end{center}
	\label{tab:sim}
\end{table}

@@

The optimal statistic was computed using the PTA data analysis package 
PAL2 %\footnote{\href{https://github.com/jellis18/PAL2}{https://github.com/jellis18/PAL2}} 
\citep{evh17a}. 
Figure~\ref{fig:os_dataset_sample} shows the fixed-noise and noise-marginalized 
optimal statistic for a single realization of the stochastic background with $\Agw=$. 
We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
for a common red process. For this particular realization of the stochastic background, 
the values of $\hat{A}^2$ from the noise-marginalized and Bayesian analyses are in good agreement 
with each other and the injected GW stochastic background, 
but the value from the fixed-noise analysis is significantly lower. 
The noise-marginalized analysis also gives a larger mean SNR compared to the fixed-noise analysis.
%\begin{figure}[ht]
%	\includegraphics[width=0.9\columnwidth]{plots/os_dataset50.pdf}
%	\caption{Optimal statistic and SNR for a simulated dataset 
%			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
%			We compare the values found using a fixed-noise analysis (dashed blue line) to the 
%			mean optimal statistic and mean SNR found by marginalizing over 
%			the individual pulsars' red noise parameters (solid orange line). 
%			We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
%			for a common red process (solid green line). 
%			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
%			For this particular realization of the stochastic background, the fixed-noise analysis 
%			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
%			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$. 
%			For comparison, the Bayesian analysis gives a mean value of $\Agw^2 = 4.5\times10^{-29}$.}
%	\label{fig:os_dataset_sample}
%\end{figure}

In Fig.~\ref{fig:os_datasetstats} we compare the fixed-noise values of $\hat{A}^2$ 
to the mean $\hat{A}^2$ for 200 different realizations of a GWB with $\Agw = 5\times10^{-15}$. 
Using the noise values from individual noise analyses 
systematically underestimates the strength of the GWB, 
while using the noise values from a common noise analysis more accurately recovers the injected value. 
The mean value of the optimal statistic from the fixed-noise analysis using the individual noise results, 
averaged over realizations of the GWB, is $\hat{A}^2 = 7.9 \times10^{-30}$ with a mean SNR of 2.3. 
The fixed-noise and noise-marginalized analyses using the common noise results both give 
$\hat{A}^2 = 2.4\times10^{-29}$ and a mean SNR of 4.1.
\begin{figure}[ht]
	\includegraphics[width=0.9\columnwidth]{plots/optstat_A5e-15.pdf}
	\caption{Optimal statistic and SNR for 200 simulated datasets 
			containing an injected GWB. 
			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
			We compare the values found by fixing the red noise 
			based on the results from individual and common noise analyses 
			(blue and orange lines, respectively) 
			to the mean optimal statistic and mean SNR found by marginalizing over 
			the red noise parameters using the results of a common noise analysis (green lines). 
			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
			that are significantly lower than the injected value, 
			while using the values from a common noise analysis more accurately estimates $\hat{A}^2$.}
	\label{fig:os_datasetstats}
\end{figure}

The fixed-noise and noise-marginalized analyses gave the same results 
for an injected GWB with $\Agw = 5\times10^{-15}$, 
but there are differences between them when analyzing datasets 
containing smaller injected values of $\Agw$. 
Figure~\ref{fig:os_datasetstats2} shows the mean optimal statistic and mean SNR 
for 100 realizations of a GWB with $\Agw = 10^{-15}$. 
The fixed-noise analysis using the individual noise values 
gives a mean value of $\hat{A}^2 = 2.9\times10^{-31}$ and a mean SNR of 1.1. 
The fixed-noise analysis using the common noise values 
gives a mean value of $\hat{A}^2 = 8.4\times10^{-31}$ and a mean SNR of 2.1, 
which the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.  
The difference between the fixed-noise and noise-marginalized analyses using the common noise values 
is very small compared to the large spread in $\hat{A}^2$ and the SNR over different realizations of the GWB.
\begin{figure}[ht]
	\includegraphics[width=0.9\columnwidth]{plots/optstat_A1e-15.pdf}
	\caption{Optimal statistic and SNR for 100 simulated datasets 
			containing an injected GWB. 
			The dashed vertical line indicates the injected value, $\hat{A}^2 = 10^{-30}$. 
			We compare the values found by fixing the red noise 
			based on the results from individual and common noise analyses 
			(blue and orange lines, respectively) 
			to the mean optimal statistic and mean SNR found by marginalizing over 
			the red noise parameters using the results of a common noise analysis (green lines). 
			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
			that are significantly lower than the injected value, 
			while using the values from a common noise analysis more accurately estimates $\hat{A}^2$.}
	\label{fig:os_datasetstats2}
\end{figure}

@@

Figure~\ref{fig:os_ORF} shows the noise-marginalized 
mean value of $\hat{A}^2$ and the mean SNR 
computed assuming monopole, dipole, and quadrupole spatial correlations 
for 200 simulated data sets. Using a monopole or dipole ORF 
gives a lower value for the mean optimal statistic and mean SNR compared to the 
Hellings-Downs ORF. 
We find a noise-marginalized mean SNR above 1.0 in 97\% of our simulated data sets 
using the quadrupole ORF, and in 50\% and 68\% of our simulated data sets 
using the monopole and dipole ORFs, respectively. 
The mean SNR using the quadrupole ORF, averaged over realizations of the GWB, is 4.1, 
and we find an SNR greater than this using the monopole and dipole ORFs in just 
3\% and 3.5\% of our simulations, respectively.
\begin{figure}[htb]
	\includegraphics[width=0.9\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
	\caption{Noise-marginalized mean optimal statistic and mean SNR for 200 simulated datasets 
			containing an injected GWB with $\Agw = 5\times10^{-15}$. 
			We compare the values found using monopole (blue), dipole (orange), 
			and quadrupole (green) spatial correlations. 
			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
	\label{fig:os_ORF}
\end{figure}

@@

The optimal statistic is a computationally inexpensive 
method for assessing the significance of an isotropic stochastic GWB in PTA data. 
The optimal statistic specifically looks for the spatial correlations characteristic 
of a GWB. These analyses are computationally expensive using a full Bayesian approach, 
but with the optimal statistic the significance of different spatial correlations 
can be computed very quickly. 
This makes the optimal statistic particularly well-suited for sky scramble analyses, 
where the ORF is scrambled many times to simulate randomizing the pulsars' sky locations.

>>>>

However, the optimal statistic gives biased results 
when the pulsars have significant red noise 
due to the strong covariance between the individual red noise parameters 
and the GWB amplitude \sv{has this ever been discussed in a paper?}. 
%In this paper, we present a technique for improving the accuracy of the optimal statistic 
%by marginalizing the optimal statistic over the individual red noise parameters 
%using the posterior distributions from a full Bayesian analysis. 
%This hybrid approach produces a more accurate 
%estimate of \Agw\ and its uncertainty. 
Here we present a technique for improving the accuracy of the optimal statistic by including an additional step: 
marginalizing over the individual pulsar red noise parameters 
using the posterior distributions from a full Bayesian analysis of all the pulsars. 
This hybrid approach produces a more precise estimate of the stochastic GWB amplitude \Agw\ 
and its uncertainty, while requiring only a few minutes over the more traditional method of computation. 
%This method is more computationally intensive than previous methods 
%for computing the optimal statistic, but still requires only minutes. 
%Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
%can be used to compute the optimal statistic with any choice of the ORF, 
%allowing additional spatial correlations to be tested very quickly.
Furthermore, the same Bayesian analysis drawn upon by the noise marginalization 
can be used to compute the optimal statistic with any choice of overlap-reduction function (ORF), 
allowing for quick testing of various spatial correlations. 
This technique was used to perform the frequentist searches 
for common red signals with monopole, dipole, and Hellings-Downs spatial correlations 
in the NANOGrav 11-year dataset \citep{abb+17b}.

@@

When constructing the residuals, 
we typically fix the red noise parameters to the values that 
maximize the likelihood. 
However, this leads to a bias in the optimal statistic and SNR because 
the individual red noise parameters and the common red noise parameters 
are highly covariant. 
In this section, we compare three different techniques for computing the 
optimal statistic for PTAs with significant red noise. 
First, we fix the individual pulsars' red noise parameters to the 
maximum-likelihood values from individual pulsar analyses. 
Second, we fix the individual pulsars' red noise parameters to the 
values that jointly maximize the likelihood for a common analysis 
of all of the pulsars in our PTA 
that searches over the pulsars' red noise parameters and 
a common red process. 
For the third method, we draw values of the individual pulsars' 
red noise parameters from the posteriors generated by a Bayesian analysis 
that includes both individual red noise and a common red noise process, 
allowing us to marginalize over the individual red noise. 
We used the PTA data analysis package 
PAL2 %\footnote{\href{https://github.com/jellis18/PAL2}{https://github.com/jellis18/PAL2}} 
\citep{evh17a}
to perform the noise analyses and compute the optimal statistic.

We used these three techniques to compute the optimal statistic 
for simulated ``NANOGrav-like'' data sets. 
These consisted of 18 MSPs with observation times up to 11 years 
and sky positions matching the 18 longest-observed pulsars in the 
NANOGrav 11-year data set \citep{abb+17}.
with the observation times increased so that the longest-observed pulsars 
had TOAs spanning over 11 years. 
We include white noise for all pulsars, plus 
red noise parametrized as a power law, $P(f) = A_\mathrm{red} f^{-\gamma}$, 
for those pulsars that show evidence of red noise 
(see Table~\ref{tab:sim} for more details).
\begin{table}[t]
	\setlength{\tabcolsep}{5pt}
	\caption{Pulsar parameters used in simulated PTAs. \sv{finish filling in this table}}
	\begin{center}
	\begin{tabular}{ccccc}
		\hline\hline
    		Pulsar	& $T_\mathrm{obs}$ (yrs) & $\sigma_w$ ($\mu\mathrm{s}$) & $A_\mathrm{red}$ & $\gamma_\mathrm{red}$ \\
		\hline
		J0030+0451 & 11.0 & & -13.93 & 3.56 \\
		J0613$-$0200 & 11.0 &  & -13.14 & 1.22 \\
		J1012+5307 & 11.0 &  & -12.79 & 1.51 \\
		J1024$-$0719 & 6.0 & & $-$ & $-$ \\
		J1455$-$3330 & 11.0 & & $-$ & $-$ \\
		J1600$-$3053 & 8.0 & & $-$ & $-$ \\
		J1614$-$2230 & 7.0 & & $-$ & $-$ \\
		J1640+2224 & 11.0 & & $-$ & $-$ \\
		J1713+0747 & 11.0 & & -14.14 & 1.58 \\
		J1741+1351 & 6.0 & & $-$ & $-$ \\
		J1744$-$1134 & 11.0 & & $-$ & $-$ \\
		B1855+09 & 11.0 & & -13.75 & 3.54 \\
		J1853+1303 & 7.0 & & $-$ & $-$ \\
		J1909$-$3744 & 11.0 &  & -13.84 & 1.74  \\
		J1918$-$0642 & 11.0 & & $-$ & $-$ \\
		J2010$-$1323 & 7.0 & & $-$ & $-$ \\
		J2145$-$0750 & 11.0 &  & -12.69 & 1.30 \\
		J2317+1439 & 11.0 & & $-$ & $-$ \\
    		\hline\hline
	\end{tabular}
	\end{center}
	\label{tab:sim}
\end{table}

@@

Figure~\ref{fig:os_dataset_sample} shows the fixed-noise and noise-marginalized 
optimal statistic for a simulation with a GWB with $\Agw=5\times10^{-15}$. 
%We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
%for a common red process. 
For this particular realization of the GWB, 
the fixed-noise analysis 
using the individual noise results gives 
$\hat{A}^2 = 6.6\times10^{-30}$ with $\mathrm{SNR} = 2.4$, 
and the fixed-noise analysis using the common noise results gives 
$\hat{A}^2 = 2.6\times10^{-29}$ with $\mathrm{SNR} = 6.0$. 
The noise-marginalized analysis gives $\hat{A}^2 = (2.5 \pm 0.1)\times10^{-29}$ 
with $\mathrm{SNR}=4.8\pm0.8$. 
The value of $\hat{A}^2$ from the fixed-noise analysis using the individual noise results 
is significantly lower than the injected level of the GWB, while 
the values of $\hat{A}^2$ from the fixed-noise analysis using the common noise results 
and the noise-marginalized analysis are in good agreement with each other and the injected value. 
The fixed-noise analysis using the individual noise results also gives a significantly lower SNR 
than the analyses using the common noise results. 
The mean SNR from the noise-marginalized analysis is smaller 
than the SNR found with the fixed-noise analysis using the common noise results, 
but the two are consistent at $1.6\sigma$.
\begin{figure}[t]
	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15_dataset11.pdf}
	\caption{Optimal statistic for a simulated dataset containing an injected GWB. 
			We compare the values found using a fixed-noise analysis with noise parameters taken from 
			individual and common noise analyses (dashed blue lines and solid orange lines, respectively) 
			to the values found by marginalizing over 
			the individual pulsars' red noise parameters (solid green lines). 
			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
			For this particular realization of the GWB, the fixed-noise analysis 
			using the individual noise results underestimates the strength of the GWB, 
			while the fixed-noise analysis using the common noise results 
			and the noise-marginalized analysis more accurately find $\hat{A}^2$.}
	\label{fig:os_dataset_sample}
\end{figure}
%\begin{figure}[ht]
%	\includegraphics[width=0.9\columnwidth]{plots/os_dataset50.pdf}
%	\caption{Optimal statistic and SNR for a simulated dataset 
%			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
%			We compare the values found using a fixed-noise analysis (dashed blue line) to the 
%			mean optimal statistic and mean SNR found by marginalizing over 
%			the individual pulsars' red noise parameters (solid orange line). 
%			We also include the distribution of $\Agw^2$ found using the results of a Bayesian analysis 
%			for a common red process (solid green line). 
%			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
%			For this particular realization of the stochastic background, the fixed-noise analysis 
%			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
%			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$. 
%			For comparison, the Bayesian analysis gives a mean value of $\Agw^2 = 4.5\times10^{-29}$.}
%	\label{fig:os_dataset_sample}
%\end{figure}

In Fig.~\ref{fig:os_datasetstats} we show the optimal statistic 
for 200 different realizations of a GWB with $\Agw = 5\times10^{-15}$ 
computed using the three techniques described above. 
Using the noise values from individual noise analyses 
systematically underestimates the strength of the GWB, 
while using the noise values from a common noise analysis more accurately recovers the injected value. 
The mean value of the optimal statistic from the fixed-noise analysis using the individual noise results, 
averaged over realizations of the GWB, is $\hat{A}^2 = 7.9 \times10^{-30}$ with a mean SNR of 2.3. 
The fixed-noise and noise-marginalized analyses using the common noise results both give 
a mean value of the optimal statistic of $\hat{A}^2 = 2.4\times10^{-29}$ and a mean SNR of 4.1.
\begin{figure}[t]
	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15.pdf}
	\caption{Optimal statistic and SNR for 200 simulated datasets 
			containing an injected GWB. 
			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
			We show the values found by fixing the red noise 
			based on the results from individual and common noise analyses 
			(blue and orange lines, respectively) 
			as well as the mean optimal statistic and mean SNR found by marginalizing over 
			the red noise parameters using the results of a common noise analysis (green lines). 
			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
			that are significantly lower than the injected value, 
			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
			that are closer to the injected value.}
	\label{fig:os_datasetstats}
\end{figure}

The fixed-noise and noise-marginalized analyses give the same results 
for an injected GWB with $\Agw = 5\times10^{-15}$, 
but there are differences between them when analyzing datasets 
containing smaller injected values of $\Agw$. 
Figure~\ref{fig:os_datasetstats2} shows the mean optimal statistic and mean SNR 
for 100 realizations of a GWB with $\Agw = 10^{-15}$. 
The fixed-noise analysis using the individual noise values 
gives a mean value of $\hat{A}^2 = 2.9\times10^{-31}$ and a mean SNR of 1.1. 
The fixed-noise analysis using the common noise values 
gives a mean value of $\hat{A}^2 = 8.4\times10^{-31}$ and a mean SNR of 2.1, 
which the noise-marginalized analysis gives a mean value of 
$\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.
\begin{figure}[t]
	\includegraphics[width=0.95\columnwidth]{plots/optstat_A1e-15.pdf}
	\caption{Optimal statistic and SNR for 100 simulated datasets 
			containing an injected GWB with $\hat{A}^2 = 10^{-30}$ (dashed vertical line). 
			We show the values found by fixing the red noise 
			based on the results from individual and common noise analyses 
			(blue and orange lines, respectively) 
			to the mean optimal statistic and mean SNR found by marginalizing over 
			the red noise parameters using the results of a common noise analysis (green lines). 
			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
			that are significantly lower than the injected value, 
			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
			that are closer to the injected value.}
	\label{fig:os_datasetstats2}
\end{figure}

We also compare these three methods 
based on how far the computed values of the optimal statistic are from the injected value of of the GWB 
compared to the uncertainty in the optimal statistic,
\begin{equation}
	s = \frac{\hat{A}^2 - \Agw^2}{\sigma} = \frac{\hat{A}^2 - \Agw^2}{\hat{A}^2/\rho} \,.
\end{equation}
Figure~\ref{fig:os_compare} shows the distribution of $(\hat{A}^2-\Agw^2)/\sigma$

@@

Figure~\ref{fig:os_ORF} shows the noise-marginalized 
mean value of $\hat{A}^2$ and the mean SNR 
computed assuming monopole, dipole, and quadrupole spatial correlations 
for 200 simulated data sets. Using a monopole or dipole ORF 
gives a lower value for the mean optimal statistic and mean SNR compared to the 
Hellings-Downs ORF. 
We find a noise-marginalized mean SNR above 1.0 in 97\% of our simulated data sets 
using the quadrupole ORF, and in 50\% and 68\% of our simulated data sets 
using the monopole and dipole ORFs, respectively. 
The mean SNR using the quadrupole ORF, averaged over realizations of the GWB, is 4.1, 
and we find an SNR greater than this using the monopole and dipole ORFs in just 
3\% and 3.5\% of our simulations, respectively.
\begin{figure}[t]
	\includegraphics[width=0.95\columnwidth]{plots/optstat_spatial_A5e-15.pdf}
	\caption{Noise-marginalized mean optimal statistic and mean SNR for 200 simulated datasets 
			containing an injected GWB with $\Agw = 5\times10^{-15}$. 
			We compare the values found using monopole (blue), dipole (orange), 
			and quadrupole (green) spatial correlations. 
			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$.}
	\label{fig:os_ORF}
\end{figure}

@@

The optimal statistic is a frequentist estimator for the strength of 
an isotropic stochastic GWB in PTA data that searches 
for a common signal with a particular spatial correlation relation 
using a matched filter approach. 
Searching for the spatial correlations characteristic of a GWB 
using a full Bayesian approach is computationally inexpensive, 
whereas the optimal statistic can be computed in a fraction of the time. 
This makes the optimal statistic particularly useful for simulations 
or for analyses where many spatial correlations are compared 
such as sky scramble analyses.
